---
title: "Instruction Manual"
author: "Marieke Dirksen"
date: "July 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction


## Course schedule

| Time | Description |
|------|-------------|
| 09:00-09:15 | Welkom & Koffie |
| 09:15-09:30 | Introduction | 
| 09:30-10:30 | Git & GitHub |
| 10:30-10:45 | Koffie break |
| 10:45-12:00 | R projects & documentation of code |
| 12:00-13:00 | Lunch Break |
| 13:00-14:30 | Putting it all together: combining Rstudio with Github |
| 14:30-14:45 | Koffie Break |
| 14:45-16:00 | Protect your code: Travis CI |
| 16:00-17:00 | Round up practicals & time for questions |
| 17:00-17:30 | Drinks |


What we are going to do in this course

* location of the course
* course schedule
* who is doing which part of the course  

The code for this course can be found [here](https://github.com/MariekeDirk/Reproducability-Reusability-course.git)

# Git and GitHub
Background information on git and github, written by the person teaching this topic

# Documentation of code in R
examples of code documentation, written by the person teaching this topic

```{r function}
#'
#'@title Square area
#'@description Function to calculate the area of a square. Requires numeric input. 
#'@param length The length of the square [m]
#'@param width The width of the square [m]
#'@examples
#'square_area(4,5)
#'
#'\dontrun{
#'square_area(4,"5")}
#'@author Marieke Dirksen
#'@export
#'

square_area<-function(length,width){
  if(!inherits(length,"numeric") | !inherits(width,"numeric")) {
    message("Your length or width is not numeric")
    return(FALSE)
  } else {
    area<-length*width
    return(area)
  }
    
}

square_area(3,4)
```

# Combining Rstudio and GitHub
examples version control in Rstudio, written by the person teaching this topic

# Travis CI
how to put your code online, written by the person teaching this topic