---
title: "Working with R packages"
author: "Marieke Dirksen"
date: "January 11, 2018"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#themes: “default”, “cerulean”, “journal”, “flatly”, “readable”, “spacelab”, “united”, “cosmo”, “lumen”, “paper”, “sandstone”, “simplex”, “yeti"
```


# Introduction
During this tutorial the following topics are covered:

* Create a new project
* Writing and documenting functions in R
* Raw data and data description
* Document, built and reload 
* Extra: Introduction to markdown, README.md

Required packages for this tutorial:

* devtools
* roxygen2

# Creating a new project
According to [Rpackages by Hadley Wickham](http://r-pkgs.had.co.nz) the following package components are included:

* Code (R/)
* Package metadata (DESCRIPTION)
* Object documentation (man/)
* Vignettes (vignettes/)
* Testing (tests/)
* Namespaces (NAMESPACE)
* Data (data/)
* Compiled code (src/)
* Installed files (inst/)
* Other components 

> Challenge 2.1: Create a new Rstudio project on GitHub ([online instructions](http://jennybc.github.io/2014-05-12-ubc/ubc-r/session2.4_github.html)). What are the standard folders and files that are created?

> Challenge 2.2: Give your new package a title, description and author name. Commit your changes to GitHub. 

>> Additional reading: We won't cover all the package components, feel free to read more about them [online ](http://r-pkgs.had.co.nz), or check the [online support](https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects)

# Writing and documenting functions in R

We use functions all the time, e.g. `min`, `max`, `mean`.

> Challenge 3.1: Check the code with `read.csv`. On which function does `read.csv` relay? 

Using function can significantly reduce your code length and improve the reproducibility. Below an example of a function in R:

```{r}
area_of_square<-function(length,width){
  if(!inherits(length,"numeric") | !inherits(width,"numeric")) { #check if the input is numeric
    message("Your length or width is not numeric")
    return(FALSE)                                                # if the input is not numeric FALSE is returned 
  } else {
    area<-length*width                                           #calculating the area of the square
    return(area)                                                 #returning the variable area
  }
    
}

```

> Challenge 3.2: Run `area_of_square` and test the function with random numbers. What happens if you use `area_of_square('5','5')`? Save your function in the R folder. 

The function `read.csv` has a documentation page, which we can read typing `?read.csv`. It is very handy to check what all the variables mean and what the function does. So, we are also going to create a documentation page for our function. The documentation uses the following syntax: 

```{r}
#'
#'@title Square area
#'@description Function to calculate the area of a square. Requires numeric input. 
#'@param length The length of the square 
#'@param width The width of the square 
#'@examples
#'square_area(4,5)
#'
#'\dontrun{
#'square_area(4,"5")}
#'@author Marieke Dirksen
#'@export
#'

area_of_square<-function(length,width){
  if(!inherits(length,"numeric") | !inherits(width,"numeric")) { #check if the input is numeric
    message("Your length or width is not numeric")
    return(FALSE)                                                # if the input is not numeric FALSE is returned 
  } else {
    area<-length*width                                           #calculating the area of the square
    return(area)                                                 #returning the variable area
  }
    
}

```

> Challenge 4.2: Write and document your own function. In order to create the documentation page use `devtools::document()`. Check your man/ folder, it should now contain a *.Rd file. Use the Install and Restart button under Build to attach the library. To test if your documentation works type: `?area_of_square`.  

>> Additional material: watch [youtube](https://www.youtube.com/watch?v=WK3_JAPP7ZM)

# Raw data and data description

Small datasets, like `data("iris")`, can be stored within a package. Similar to the functions this data is also documented: `?iris`. We are going to create our own dataset with lengths and widths from which we want to calculate the area. 

```{r data}
length<-runif(15,min=1,max=50)
width<-runif(15,min=1,max=50)
my.df<-data.frame("length"=length,"width"=width)

devtools::use_data(my.df,overwrite=TRUE)
```

> Challenge 5.1: Use `devtools::use_data_raw()` to create the folder data-raw. Go to the folder and create a R scipt with the name dataraw.R. Create a `data.frame` for your own function, use `devtools::use_data(my.df)` to write your data into the data folder. 

The documentation of the data looks similar to the function documentation:

```{r data documentation}
#' Lengths and widths
#'
#' This dataset contains random numerical values between 1 and 50
#' 
#'
#' @format A data frame with 15 rows and 2 variables:
#' \describe{
#'   \item{length}{length of the area}
#'   \item{width}{width of the area}
#' }
"my.df"

```

> Challenge 5.2: Write the documentation of your own function and add it to your package. 

# Extra: Introduction to markdown, README.md

It is a good practice to keep a README file. A fancy extention nowadays is '.md'. 

> Challenge 7.1: Look at the example [from the R's markdown package](https://github.com/rstudio/rmarkdown) and look at the [instructions](http://stat545.com/packages05_foofactors-package-02.html#use-readme.rmd). Edit the README.md file for this package.  

